\plaintitle

### 2025-03-14 docs/style/article-body.tex
linebreak
final corrections before hand-in

In computer science, there are many applications for randomly generated numbers. From generating keys for cryptography

However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}. These numbers do not generate true randomness, and in order to heighten security other methods of generating these are required. Current random number generators (\emph{RNG}) are usually implemented in code, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}. This pseudo-random number generation (\emph{PRNG}) comes with the drawback that the result is always deterministic, provided that the initial state is known.

@@

As described in Clason's work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness. The device has an LED and a photodiode positioned a few millimeters apart, ensuring efficient light coupling. The photodiode detects light from the LED, and converts the light into a current signal, which is sent to a transimpedance amplifier to convert it into measurable voltage. In order to minimize disruptions by other external lights, the system is enclosed in a shielded measurment box.

@@

Microcontrollers (MCUs) are compact and low-power computing devices designed primary for embedded systems and real-time applications. Unlike general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip. Thus, enabling efficient and autonomous operation in constrained environments. Various development frameworks and tools are available for programming MCUs, including the Arduino Framework, manufacturer specific SDKs and raw C++, among others. Each framework presents different trade-offs in terms of performance overhead, efficiency and ease of development.

@@

The Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7 is especially good for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and DSP capabilities. The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and SIMD-style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction. SIMD-controlled DSP architectures, as described by Han et al. \cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations. Thus, making them highly effective for Toeplitz matrix-vector multiplications. Although, the Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction. Moreover, its USB High-Speed (480 Mbit/sec) interface, supporting both USB device and USB host modes, allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.

@@

To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al. \cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) are multiplied with a pre-determined seed matrix (\(K\)). The size of the seed is directly dependant on the size of the sampled data, and can be fixed or continually re-sampled as needed. To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent. The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number. An example of how this extraction works can be seen below.

@@

This allows us to intepret the result as a random 4-bit integer:

@@

Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitzmatrix. Numerous systems, such as those developed by Chouhan et al. \cite{toeplitz-desc} and Zhang et al. \cite{toeplitz}, depends on fixed seeds. Fixed seeds can create security risks over prolonged operation time. To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG. Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed. Thus, minimizing temporal correlations between extractions. Furthermore, an external seed updateing mechanism via PCIe ensures that seeds are refreshed whenever a certain secutiry limit is reached. Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.

Efficient Toeplitz matrix-vector multiplication (TMVM) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments. Liao et al. \cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverser (IFFT). Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\). Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds. By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput. Thus, potentially could improve performance and reduce latency.

@@

Each iteration results in an implementation written in C, all of which will be supplied as an appendix to this paper. Furthermore, to simplify the implementation process, we will use the Arduino framework to communicate with the hardware. Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point.

@@

2025-03-12: Elaborated on background, as well as adding more details regarding hardware. Note that the hardware selected is subject to change over time. Further elaborated on related works in optimizing Toeplitz extraction.

>>>>

In computer science, there are many applications for randomly generated numbers. From generating keys for cryptography, salting password hashes, load balancing in distributed systems, memory addressing and so much more. However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}. These numbers do not generate true randomness, and in order to heighten security other methods of generating these are required. Current random number generators (\emph{RNG}) are usually implemented in code, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}. This pseudo-random number generation (\emph{PRNG}) comes with the drawback that the result is always deterministic, provided that the initial state is known.

@@

As described in Clasons work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness. The device has an LED and a photodiode positioned a few millimeters apart, ensuring efficient light coupling. The photodiode detects light from the LED, and converts the light into a current signal, which is sent to a transimpedance amplifier to convert it into measurable voltage. In order to minimize disruptions by other external lights, the system is enclosed in a shielded measurment box.

@@

Microcontrollers (MCUs) are compact and low-power computing devices designed primary for embedded systems and real-time applications. Unlike general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip. Thus, enabling efficient and autonomous operation in constrained environments. Various development frameworks and tools are available for programming MCUs, including the Arduino Framework, manufacturer specific SDKs and raw C++, among others. Each framework presents different trade-offs in terms of performance overhead, efficiency, and ease of development.

@@

The Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7 is especially good for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and DSP capabilities. The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and SIMD-style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction. SIMD-controlled DSP architectures, as described by Han et al. \cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations. Thus, making them highly effective for Toeplitz matrix-vector multiplications. The Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction. Moreover, its USB High-Speed (480 Mbit/sec) interface, supporting both USB device and USB host modes, allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.

@@

To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al. \cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) are multiplied with a pre-determined seed matrix (\(K\)). The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed. To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent. The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number. An example of how this extraction works can be seen below.

@@

This allows us to interpret the result as a random 4-bit integer:

@@

Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitz matrix. Numerous systems, such as those developed by Chouhan et al. \cite{toeplitz-desc} and Zhang et al. \cite{toeplitz}, depends on fixed seeds. Fixed seeds can create security risks over prolonged operation time. To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG. Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed. Thus, minimizing temporal correlations between extractions. Furthermore, an external seed updating mechanism via PCIe ensures that seeds are refreshed whenever a certain security limit is reached. Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.

Efficient Toeplitz matrix-vector multiplication (TMVM) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments. Liao et al. \cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverse (IFFT). Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\). Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds. By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput. Thus, potentially could improve performance and reduce latency.

@@

Each iteration results in an implementation written in C, all of which will be supplied as an appendix to this paper. Furthermore, to simplify the implementation process, we will use the Arduino framework to communicate with the hardware. Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point. \newpage

@@

2025-03-13: Elaborated on background, as well as adding more details regarding hardware. Note that the hardware selected is subject to change over time. Further elaborated on related works in optimizing Toeplitz extraction.
